---
title: "Lab_clean"
author: "Yara Yachnyk"
date: "2025-03-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Function for cross-sectional data for all datasets in period 2019-2025:

```{r}
library(dplyr)
library(lubridate)
library(fastDummies)

clean_and_combine <- function(data_list) {
  data_cleaned <- data_list %>%
    bind_rows() %>%
    distinct() %>%
    select(-any_of("adjusted_price")) %>%
    mutate(
      price = gsub("\\$", "", price),  # Remove $ sign
      price = gsub(",", "", price),    # Remove commas
      price = as.numeric(price),       # Convert to numeric
      date = as.Date(date)             # Convert to Date type
    ) %>%
    drop_na() %>%  # Drop missing values

    # Ensure correct date format before grouping
    mutate(year_month_day = format(date, "%Y-%m-%d")) %>%

    # Group and calculate average price
    group_by(year_month_day) %>%
    summarise(avg_price = mean(price, na.rm = TRUE), .groups = "drop") %>%
    
    # Add additional date-based variables
    mutate(
      month = month(as.Date(year_month_day)),  # Extract month
      year = year(as.Date(year_month_day)),    # Extract year
      is_weekend = wday(as.Date(year_month_day), week_start = 1) >= 6  # Weekend indicator
    ) %>%

    # Create dummy variables for months
    dummy_cols(select_columns = "month", remove_first_dummy = TRUE)

  return(data_cleaned)
}

```

Let's upload dataset we will be working with and combine them in proper years:

```{r}
library(readr)

# year 2024-2025
X2024_09_24 <- read_csv("data/2024-09-24.csv")
X2024_06_10 <- read_csv("data/2024-06-10.csv")
X2024_03_11 <- read_csv("data/2024-03-11.csv")

#year 2023-2024
X2023_12_12 <- read_csv("data/2023-12-12.csv")
X2023_09_03 <- read_csv("data/2023-09-03.csv")
X2023_06_05 <- read_csv("data/2023-06-05.csv")
X2023_03_09 <- read_csv("data/2023-03-09.csv")

#year 2022-2023
X2022_09_07 <- read_csv("data/2022-09-07.csv")
X2022_06_05 <- read_csv("data/2022-06-05.csv")
X2022_03_08 <- read_csv("data/2022-03-08.csv")
# datasets per each year
y_2024_2025 <- list(X2024_09_24, X2024_06_10, X2024_03_11)
cleaned_2024_2025 <- clean_and_combine(y_2024_2025)

y_2023_2024 <- list(X2023_12_12, X2023_09_03, X2023_06_05, X2023_03_09)
cleaned_2023_2024 <- clean_and_combine(y_2023_2024)

y_2022_2023 <- list(X2022_09_07, X2022_06_05, X2022_03_08)
cleaned_2022_2023 <- clean_and_combine(y_2022_2023)
```
We split them separately to not overload R software

```{r}
#year 2019-2022 in 3 separate datasets
X2021_12_05 <- read_csv("data/2021-12-05.csv")
X2021_09_07 <- read_csv("data/2021-09-07.csv")
X2021_06_03 <- read_csv("data/2021-06-03.csv")
X2021_03_04 <- read_csv("data/2021-03-04.csv")
X2021_02_10 <- read_csv("data/2021-02-10.csv")
X2021_01_10 <- read_csv("data/2021-01-10.csv")

y_2021_2022 <- list(X2021_12_05, X2021_09_07, X2021_06_03, X2021_03_04, X2021_02_10, X2021_01_10)
cleaned_2021_2022 <- clean_and_combine(y_2021_2022)

X2020_12_13 <- read_csv("data/2020-12-13.csv")
X2020_11_03 <- read_csv("data/2020-11-03.csv")
X2020_10_10 <- read_csv("data/2020-10-10.csv")
X2020_09_10 <- read_csv("data/2020-09-10.csv")
X2020_08_18 <- read_csv("data/2020-08-18.csv")
X2020_07_10 <- read_csv("data/2020-07-10.csv")
X2020_06_08 <- read_csv("data/2020-06-08.csv")
X2020_05_08 <- read_csv("data/2020-05-08.csv")
X2020_04_16 <- read_csv("data/2020-04-16.csv")
X2020_03_13 <- read_csv("data/2020-03-13.csv")
X2020_02_14 <- read_csv("data/2020-02-14.csv")
X2020_01_05 <- read_csv("data/2020-01-05.csv")

y_2020_2021 <- list(
  X2020_12_13, X2020_11_03, X2020_10_10, X2020_09_10, 
  X2020_08_18, X2020_07_10, X2020_06_08, X2020_05_08, 
  X2020_04_16, X2020_03_13, X2020_02_14, X2020_01_05
)

cleaned_2020_2021 <- clean_and_combine(y_2020_2021)

# 2019-2020
X2019_10_15 <- read_csv("data/2019-10-15.csv")
X2019_04_08 <- read_csv("data/2019-04-08.csv")
y_2019_2020 <- list(X2019_10_15, X2019_04_08)
cleaned_2019_2020 <- clean_and_combine(y_2019_2020)
```

Function for regression for every dataset

```{r}
# Run function for regression (Yara's algorithm for avg price regression taken from Dante)
run_avg <- function(cleaned) {
  reg_model <- lm(avg_price ~ month + is_weekend + month:is_weekend, data = cleaned)
  return(summary(reg_model))
}

run_avg(cleaned_2024_2025)
run_avg(cleaned_2023_2024)
run_avg(cleaned_2022_2023)
run_avg(cleaned_2021_2022)
run_avg(cleaned_2020_2021)
run_avg(cleaned_2019_2020)
```


Then let's do the same with panel data:
First we create holidays function to generate holidays across 2019-2025 years

```{r}
#generate Dutch public holidays

install.packages("timeDate")
library(dplyr)
library(lubridate)
library(plm)
library(timeDate)  # Required for calculating Easter Sunday: because this day is each year different. 

# Function to generate Dutch public holidays for a given year
dutch_holidays_for_year <- function(year) {
  easter_sunday <- as.Date(Easter(year))  # Compute Easter Sunday
  holidays <- as.Date(c(
    paste0(year, "-01-01"), # New Year's Day
    paste0(year, "-04-27"), # King's Day
    easter_sunday + 1,      # Easter Monday
    paste0(year, "-05-05"),
    paste0(year, "-12-25"), # Christmas Day
    paste0(year, "-12-26")  # Boxing Day
  ))
  return(holidays)
}

# Generate holidays for a range of years
generate_dutch_holidays <- function(start_year, end_year) {
  holidays <- unlist(lapply(start_year:end_year, dutch_holidays_for_year))
  return(as.Date(holidays))
}

# Define holiday range dynamically
dutch_holidays <- generate_dutch_holidays(2019, 2025)

```

Function for years 2019-2025:

```{r}
# Main function: Clean, combine, and process data
clean_panel_data <- function(data_list, start_year, end_year) {
  
  # Generate Dutch holidays for the given year range
  dutch_holidays <- generate_dutch_holidays(start_year, end_year)

  # Data cleaning and transformation
  combined_data <- data_list %>%
    bind_rows() %>%
    distinct() %>%
    select(-any_of("adjusted_prices")) %>%
    mutate(
      price = as.numeric(gsub("\\$", "", price)),  # Remove $ sign
      price = gsub(",", "", price),               # Remove commas
      price = as.numeric(price),                  # Convert to numeric
      date = as.Date(date),                        # Ensure date format
      year_month_day = format(date, "%Y-%m-%d"),  # Extract year-month-day
      is_weekend = wday(date, week_start = 1) >= 6, # Weekend indicator
      is_holiday = date %in% dutch_holidays        # Holiday indicator
    ) %>%
    drop_na() %>%
    mutate(
      month_num = factor(month(date)),   # Extract month and convert to factor
      year = year(date)                  # Extract year
    ) %>%
    group_by(listing_id, date, month_num, is_weekend, is_holiday) %>%
    summarise(avg_price = mean(price, na.rm = TRUE), .groups = "drop")

  # Convert to panel data format
  pdata <- pdata.frame(combined_data, index = c("listing_id", "date"))

  return(pdata)
}
```

```{r}
# Run the fnction for all the years: 
pdata_2019_2020 <- clean_panel_data(y_2019_2020, 2019, 2020)
pdata_2020_2021 <- clean_panel_data(y_2020_2021, 2020, 2021)
pdata_2021_2022 <- clean_panel_data(y_2021_2022, 2021, 2022)
pdata_2022_2023 <- clean_panel_data(y_2022_2023, 2022, 2023)
pdata_2023_2024 <- clean_panel_data(y_2023_2024, 2023, 2024)
pdata_2024_2025 <- clean_panel_data(y_2024_2025, 2024, 2025)
```

```{r}
y_2019_2025 <- bind_rows(y_2019_2020, y_2020_2021,  y_2021_2022,  y_2022_2023,  y_2023_2024,  y_2024_2025)
pdata_2019_2025 <- clean_panel_data(y_2019_2025, 2019, 2025) ## this is under question for Dante
```

Now regression for panel data:

```{r}
# Run panel regression (only needed for separate datasets, under question)
run_p_reg <- function(pdata) {
  panel_model <- plm(avg_price ~ month_num:is_weekend + month_num + is_weekend + is_holiday, 
                     data = pdata, model = "within")
  return(summary(panel_model))
}

# Example usage:
run_p_reg(pdata_2025_2024)
run_p_reg(pdata_2023_2024)
run_p_reg(pdata_2022_2023)
run_p_reg(pdata_2021_2022)
run_p_reg(pdata_2020_2021)
run_p_reg(pdata_2019_2020)
```

